### 2019-01-04 revisions/review-responses.tex

        %---------------------------------------------------------------------
        \question  Table 3: ``cycle time'' is not defined in the paper. Please 
        add.
        \begin{solution}
                        The ``cycle time'' definition has been added at the first 
                        appearance in text. 
        \end{solution}

        %---------------------------------------------------------------------
        \question  Page 14, line 224. The 3 day time step as the ``optimum'' for 
        Th fuel cycles in an MSR was first described and concluded by Powers et 
        al. Please add a reference to their initial work.
        \begin{solution}
        		The reference to \cite{powers_new_2013} has been added.
        \end{solution}

        %---------------------------------------------------------------------
        \question  Page 14, line 234 onwards: Doesn't SERPENT already have an 
        MSR removal capability? If so, what is different about using SaltProc 
        with SERPENT?
        \begin{solution}
                It does. Indeed, much of the inspiration for SaltProc was the 
                lack of documentation, difficulty of use, and various 
                inexplicable results we acquired when we tried use this 
                built-in capability. This capability has not performed well and 
                needs to be better documented and more robustly verified before 
                we can use it. We fully intend to contribute to its improvement 
                if it can be made to agree with expectations.  Additional text 
                to clarify this point has been add to solution for question 
                \ref{built-in}.                
        \end{solution}

        %---------------------------------------------------------------------
        \question  Page 18, Figure 7: The figure is hard to interpret or see 
        clearly what is going on. Could an additional figure or a zoomed in 
        portion be added to show what the swing in k is over a much shorter 
        time interval? k seems to be swinging dramatically but over what time 
        period and how would this be controlled in reality? The graph almost 
        suggests that the core is unstable??
        \begin{solution}
                Zoomed portion for 150 EFPD interval has been added. We also 
                added notes on a plot to explain swing in multiplication factor.
        \end{solution}

@@

        %--------------------------------------------------------------------
        \question  How did the authors verify the coupled SaltProc and Serpent 
        code system?  
        \begin{solution}
                We have compared few parameters (multiplication factor, Th refill 
                rate, neutron energy spectrum) with Betzler \emph{et al.} 										
                \cite{betzler_molten_2017} and mentioned it in the Result section. 
                We also compared neutron energy spectrum and temperature coefficients
                 for equilibrium composition with Park \emph{et al.} \cite{park_whole_2015}.
                In a future SaltProc release suite of unit tests will be added 
                to make sure that results are consistent with our test cases.
        \end{solution}

        %--------------------------------------------------------------------
        \question  In Page 3, the title of ``Table 1'' should not only contain 
        the ``fast spectrum system''. The work published by Zhou and Yang et al. 
        with the title of ``Fuel cycle analysis of molten salt reactors based on 
        coupled neutronics and thermal-hydraulics calculations'' needs to be 
        included in Table 1 for completeness.
        \begin{solution}
                Thank you for your excellent recommendations. The table has been 
                enriched as requested.
        \end{solution}

        %--------------------------------------------------------------------
        \question  In Page 5, the following sentence needs to be explained. ``We 
        employed this extended SERPENT 2 for a simplified unit-cell geometry of 
        thermal spectrum thorium-fueled MSBR and obtained results which 
        contradict existing MSBR depletion simulations.''
        \begin{solution}
                This statement has been significantly extended (see solution 
                \ref{built-in}).
        \end{solution}

### 2019-01-04 revisions/review-responses.tex

        %---------------------------------------------------------------------
        \question  Page 20, Fig 9: What are the wiggles and dips , especially 
        seen for Np235?
        \begin{solution}
		        Explanation of this phenomenon has been added as follows:

                          ``Small dips in neptunium and plutonium number density 
                          every 16 years are caused by removing $^{237}$Np and 
                          $^{242}$Pu (included in Processing group ``Higher 
                          nuclides'', see Table~\ref{tab:reprocessing_list}) 
                          which decays into $^{235}$Np and $^{239}$Pu, 
                          respectively.''
        \end{solution}

        %---------------------------------------------------------------------
        \question  Page 20, line 351: It is more than just the Pu isotopes that 
        makes the spectrum harder? What about the other MAs etc?
        \begin{solution}
        		Thank you for the excellent point. The corrected sentence reads thus:

                         ``The neutron energy spectrum at equilibrium is harder 
                         than at startup due to plutonium and other strong 
                         absorbers accumulating in the core during reactor 
                         operation.''
        \end{solution}

### 2019-01-04 abstract.tex

Today, climate change drives humanity's search for new ways to generate 
carbon-free, reliable base-load power. Thus,
interest in advanced nuclear energy and particularly \glspl{MSR} has resurged, 
with multiple new companies pursuing commercialization of \gls{MSR} 
designs. To further 
develop these \gls{MSR} concepts, researchers need simulation tools for analyzing liquid 
fueled \gls{MSR} depletion and fuel processing. However, most 
contemporary nuclear reactor physics software is unable to perform high-fidelity 
full-core depletion calculations for a reactor design with online reprocessing. This paper 
introduces a Python package, SaltProc, which couples with the Monte Carlo code, 
SERPENT2, to simulate \gls{MSR} online reprocessing by modeling the changing 
isotopic composition of \gls{MSR} fuel salt. This work demonstrates SaltProc capabilities
for a full-core, high-fidelity model of the commercial \gls{MSBR} 
concept and verifies these results to results in the literature from 
independent, lower-fidelity analyses.

>>>>

In the search for new ways to generate carbon-free, reliable base-load power, 
interest in advanced nuclear energy and particularly \glspl{MSR} has resurged, 
with multiple new companies pursuing commercialization of \gls{MSR} 
designs. To further 
develop these \gls{MSR} concepts, researchers need simulation tools for analyzing liquid 
fueled \gls{MSR} depletion and fuel processing. However, most 
contemporary nuclear reactor physics software is unable to perform high-fidelity 
full-core depletion calculations for a reactor design with online reprocessing. This paper 
introduces a Python package, SaltProc, which couples with the Monte Carlo code, 
SERPENT2 to simulate \gls{MSR} online reprocessing by modeling the changing 
isotopic composition of \gls{MSR} fuel salt. This work demonstrates SaltProc capabilities
for a full-core, high-fidelity model of the commercial \gls{MSBR} 
concept and verifies these results to results in the literature from 
independent, lower-fidelity analyses.

### 2019-01-04 introduction.tex

The works described in \cite{park_whole_2015} and \cite{jeong_equilibrium_2016} 
are most similar to the work presented in this paper. However, few major differences 
worse to be mentioned: (1) Park \emph{et al.} employed MCNP6 for depletion 
simulations while SERPENT2 was used in this work (SERPENT2 seems to be much faster in 
massively parallel simulations involving more than 1000 cores); (2) full-core reactor 
model herein is more detailed and has no major approximation comparing with Park 
\emph{et al.} (detailed full-core models comparison might be found elsewhere
\cite{rykhlevskii_full-core_2017}); (3) Park \emph{et al.} and Jeong \emph{et al.} 
both considered volatile gases, noble metals removal and $^{233}$Pa separation while 
in the current work we are implemented detailed reprocessing scheme described in 
conceptual \gls{MSBR} design \cite{robertson_conceptual_1971}; (4) $^{232}$Th neutron 
capture reaction rate has been investigated to prove advantages of two-region core 
design; (5) the effect of removing fission product removal from fuel salt has been 
studied.

>>>>

The works described in \cite{park_whole_2015} and \cite{jeong_equilibrium_2016} 
are most similar to the work presented in this paper. However, a few major 
differences follow: (1) Park \emph{et al.} employed MCNP6 for depletion 
simulations while this work used SERPENT2; (2) the full-core reactor 
geometry herein is more detailed \cite{rykhlevskii_full-core_2017}; (3) Park \emph{et al.} and Jeong \emph{et al.} 
both only considered volatile gas removal, noble metal removal, and $^{233}$Pa separation while 
the current work implemented the more detailed reprocessing scheme specified in 
the conceptual \gls{MSBR} design \cite{robertson_conceptual_1971}; (4) the $^{232}$Th neutron 
capture reaction rate has been investigated to prove advantages of two-region core 
design; (5) the current work explicitly examines the independent impacts of removing specific fission product groups.

### 2019-01-04 results.tex

\subsection{Neutron spectrum}
Figure~\ref{fig:spectrum} shows the normalized neutron flux spectrum for the 
full-core \gls{MSBR} model in the energy range from $10^{-8}$ to $10$ MeV. The 
neutron energy spectrum at equilibrium is harder than at startup due to 
plutonium and other strong absorbers accumulation in the core during reactor 
operation.  
\begin{figure}[ht!] % replace 't' with 'b'         to force it to 
  \centering
  \includegraphics[width=\textwidth]{spectrum.png} \caption{The neutron flux energy 
  spectrum is normalized by unit lethargy and the area under the curve is normalized to 1 for initial and equilibrium fuel salt 
  composition.}
  \label{fig:spectrum}
\end{figure}

@@

\begin{tabularx}{\textwidth}{ m | s | s | x } \hline
   Reactivity coefficient [pcm/K]  & Initial      & Equilibrium  & Reference \qquad (initial)\cite{robertson_conceptual_1971} \\  \hline

@@

Table~\ref{tab:six_factor} summarizes the six factors for both initial and 
equilibrium fuel salt composition. Using SERPENT2 and SaltProc, tese factors and their statistical uncertainties
 have been calculated for both initial and equilibrium fuel salt composition (see 
Table~\ref{tab:msbr_tab}). The fast and thermal non-leakage probabilities 
remain constant despite the evolving neutron spectrum during operation. In 
contrast, the neutron reproduction factor ($\eta$), resonance escape 
probability ($p$), and fast fission factor ($\epsilon$) are considerably different between startup and 
equilibrium. As indicated in Figure~\ref{fig:spectrum}, the neutron spectrum is 
softer at the beginning of reactor life. Neutron spectrum hardening causes the fast 
fission factor to increase through the core lifetime. The opposite is true for the 
resonance escape probability. Finally, the neutron reproduction factor 
decreases during reactor operation due to accumulation of fissile plutonium 
isotopes.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[hb!]
  \centering
  \caption{Six factors for the full-core \gls{MSBR} model for initial and 
  equilibrium fuel composition.}
\begin{tabularx}{\textwidth}{ b | s | s } \hline
Factor  & Initial      & Equilibrium   \\ \hline
Neutron reproduction factor ($\eta$)     & $1.3960\pm.000052$     & 
        $1.3778\pm.00005$ \\ Thermal utilization factor (f)           & 
        $0.9670\pm.000011$     & $0.9706\pm.00001$ \\
Resonance escape probability (p)         & $0.6044\pm.000039$     & 
        $0.5761\pm.00004$ \\
Fast fission factor ($\epsilon$)         & $1.3421\pm.000040$     & 
        $1.3609\pm.00004$ \\
Fast non-leakage probability (P$_f$)     & $0.9999\pm.000004$     & 
        $0.9999\pm.000004$ \\
Thermal non-leakage probability (P$_t$)  & $0.9894\pm.000005$     & 
        $0.9912\pm.00005$ \\ \hline
\end{tabularx}
  \label{tab:six_factor}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Thorium refill rate}
In a \gls{MSBR} reprocessing scheme, the only external feed material flow  is 
$^{232}$Th. Figure~\ref{fig:th_refill} shows the $^{232}$Th feed rate 
calculated for 60 years of reactor operation. The $^{232}$Th feed rate 
fluctuates significantly as a result of the batch-wise nature of this online 
reprocessing approach. For example, the large spikes of up to 36 kg/day in a 
thorium consumption occurs every 3435 days. This is required due to batch-wise 
removal of strong absorbers (Rb, Sr, Cs, Ba). The corresponding effective 
multiplication factor increase (Figure~\ref{fig:keff}) and breeding 
intensification leads to additional $^{232}$Th consumption.  
\begin{figure}[ht!] % replace 't' with 'b' to force it to \centering
  \includegraphics[width=\textwidth]{Th_refill_rate.png} \caption{$^{232}$Th 
  feed rate over 60 years of \gls{MSBR} operation.}
  \label{fig:th_refill}
\end{figure}
\begin{figure}[ht!] % replace 't' with 'b' to force it to \centering
  \includegraphics[width=\textwidth]{Th_refill_rate_zoomed.png} \caption{Zoomed $^{232}$Th 
  feed rate for 150-EFPD time interval.}
  \label{fig:th_refill}
\end{figure}

>>>>

\subsection{Neutron spectrum}
Figure~\ref{fig:spectrum} shows the normalized neutron flux spectrum for the 
full-core \gls{MSBR} model in the energy range from $10^{-8}$ to $10$ MeV. The 
neutron energy spectrum at equilibrium is harder than at startup due to 
plutonium and other strong absorbers accumulating in the core during reactor 
operation.  
\begin{figure}[ht!] % replace 't' with 'b'         to force it to 
  \centering
  \includegraphics[width=\textwidth]{spectrum.png} \caption{The neutron flux energy 
  spectrum is normalized by unit lethargy and the area under the curve is normalized to 1 for initial and equilibrium fuel salt 
  composition.}
  \label{fig:spectrum}
\end{figure}

@@

\begin{tabularx}{\textwidth}{ X | r | r | r } \hline
Reactivity coefficient               & Initial         & Equilibrium     & Reference                                 \\ \hline
                                        & [pcm/k]         &  [pcm/k]        & (initial)\cite{robertson_conceptual_1971} \\  \hline

@@

Table~\ref{tab:six_factor} summarizes the six factors for both initial and 
equilibrium fuel salt composition. Using SERPENT2 and SaltProc, tese factors and their statistical uncertainties
 have been calculated for both initial and equilibrium fuel salt composition (see 
Table~\ref{tab:msbr_tab}). The fast and thermal non-leakage probabilities 
remain constant despite the evolving neutron spectrum during operation. In 
contrast, the neutron reproduction factor ($\eta$), resonance escape 
probability ($p$), and fast fission factor ($\epsilon$) are considerably different between startup and 
equilibrium. As indicated in Figure~\ref{fig:spectrum}, the neutron spectrum is 
softer at the beginning of reactor life. Neutron spectrum hardening causes the fast 
fission factor to increase through the core lifetime. The opposite is true for the 
resonance escape probability. Finally, the neutron reproduction factor 
decreases during reactor operation due to accumulation of fissile plutonium 
isotopes.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[hb!]
  \centering
  \caption{Six factors for the full-core \gls{MSBR} model for initial and 
  equilibrium fuel composition.}
\begin{tabularx}{\textwidth}{ b | s | s } \hline
Factor  & Initial      & Equilibrium   \\ \hline
Neutron reproduction factor ($\eta$)     & $1.3960\pm.000052$     & 
        $1.3778\pm.00005$ \\ Thermal utilization factor (f)           & 
        $0.9670\pm.000011$     & $0.9706\pm.00001$ \\
Resonance escape probability (p)         & $0.6044\pm.000039$     & 
        $0.5761\pm.00004$ \\
Fast fission factor ($\epsilon$)         & $1.3421\pm.000040$     & 
        $1.3609\pm.00004$ \\
Fast non-leakage probability (P$_f$)     & $0.9999\pm.000004$     & 
        $0.9999\pm.000004$ \\
Thermal non-leakage probability (P$_t$)  & $0.9894\pm.000005$     & 
        $0.9912\pm.00005$ \\ \hline
\end{tabularx}
  \label{tab:six_factor}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Thorium refill rate}
In a \gls{MSBR} reprocessing scheme, the only external feed material flow  is 
$^{232}$Th. Figure~\ref{fig:th_refill} shows the $^{232}$Th feed rate 
calculated for 60 years of reactor operation. The $^{232}$Th feed rate 
fluctuates significantly as a result of the batch-wise nature of this online 
reprocessing approach. Figure~\ref{fig:th_refill_zoomed} shows zoomed thorium feed 
rate for short 150-EFPD interval. Note that the large spikes of up to 36 kg/day in a 
thorium consumption occurs every 3435 days. This is required due to strong absorbers 
(Rb, Sr, Cs, Ba) removal at the end of effective cycle (100\% of these elements 
removing every 3435 days of operation). The corresponding effective 
multiplication factor increase (Figure~\ref{fig:keff}) and breeding 
intensification leads to additional $^{232}$Th consumption.  
\begin{figure}[ht!] % replace 't' with 'b' to force it to \centering
  \includegraphics[width=\textwidth]{Th_refill_rate.png} \caption{$^{232}$Th 
  feed rate over 60 years of \gls{MSBR} operation.}
  \label{fig:th_refill}
\end{figure}
\begin{figure}[ht!] % replace 't' with 'b' to force it to \centering
  \includegraphics[width=\textwidth]{Th_refill_rate_zoomed.png} \caption{Zoomed $^{232}$Th 
  feed rate for 150-EFPD time interval.}
  \label{fig:th_refill_zoomed}
\end{figure}
